Predicting sales of newly published books

### 2015-08-19 paper.tex
traducido el texto en español + arreglada una traducción en el archivo de traduccion de nombres de variables

En el caso de libros ya publicados, se dispone de datos históricos que ayudan a la generación de modelos predictivos. 
Sin embargo, en el caso de nuevos libros no hay datos históricos que ayuden a definir patrones, y sólo ciertos datos descriptivos están disponibles, por lo que no todos los métodos de predicción se pueden utilizar \cite{Madsen2008,ChernWSF15}.

@@

A lo largo de esta investigación se colaboró con la empresa Trevenque Editorial \footnote{{\tt http://editorial.trevenque.es}}, una compañía que ofrece sistemas de gestión y servicios para librerías, editoriales y distribuidoras.

La empresa Trevenque, cuyo experto asesoró esta investigación, facilitó un conjunto de datos inicial consistente en 3159 libros. 
En este conjunto, para cada libro se incluye información descriptiva del libro, como son el precio, lugares donde se han puesto a la venta, materia, editorial, ventas, devoluciones, entre otros datos. 
Table \ref{tabla:paramsOrig50} shows the original set of parameters describing each book.

@@

\begin{minipage}{2.3in} \begin{verbatim}
REFERENCE
AUTHOR
RETAIL PRICE
SUBJECT-MATTER1
SUBJECT-MATTER2
SUBJECT-MATTER3
EDITORIAL
COLLECTION
BOOKBINDING
PRINT-RUN
TOTAL SALES
MALLS SALES
DELEGATES SALES
REST OF SALES
TOTAL SALES FIRST YEAR
MALL SALES FIRST YEAR
DELEGATES SALES FIRST YEAR
REST OF SALES FIRST YEAR
TOTAL DISTRIBUTED
MALLS DISTRIBUTED
DELEGATES DISTRIBUTED
REST OF DISTRIBUTED
TOTAL DISTRIBUTED FIRST YEAR
MALLS DISTRIBUTED FIRST YEAR
DELEGATES DISTRIBUTED FIRST YEAR
\end{verbatim} \end{minipage}     &

@@

En el problema de la predicción de ventas de nuevos libros afectan diversas variables propias del libro en cuestión, como por ejemplo el autor, el precio, la materia o la editorial.

Sin embargo, existen otras variables externas que también afectan a las ventas de un nuevo libro.
Se trata de cuestiones como la estacionalidad, modas o promociones.
Estas otras variables no medibles y que los métodos de predicción clásicos no pueden procesar y que un experto debe tener en cuenta modificando la predicción ofrecida por el método automático. 
Las empresas editoriales suelen finalmente definir reglas o índices para ajustar subjetivamente los resultados de predicción.

@@

Hay varias dificultades inherentes al problema de la predicción de ventas de nuevos productos: la limitación en la cantidad de datos disponibles, la aplicación de los diferentes métodos y la elección del más adecuado de ellos.

Además, no existe una metodología estándar para resolver el problema general de la predicción de ventas. 
Habitualmente es necesario un análisis previo de los datos para eliminar las variables que describen el producto y que no aportan información útil para realizar el proceso de predicción.

@@

Thus, in this study there are three main stages (see Figure \ref{fig:stages} for details): 
\begin{itemize}
  \item  First, after the editorial company has provided the raw data, taken directly from its database, a selection variables process is carried out. In this stage, an expert sales manager, using the information obtained analyzing the SOM cluster method results, removes some unrelated variables and selects the best variables to be considered in the forecasted model. This task may help to identify leading indicators to be used by the forecasting methods in order to improve the accuracy of future forecasts.

@@

  \item \emph{Root relative squared error} (RRSE):
        \begin{equation}\label{eq:RRSE}
            % RRSE = \frac{ \sqrt{ \frac{1}{n}\sum_{i=1}^n {(p_i - o_i)}^2 } }{ \sqrt{ \frac{1}{n}\sum_{i=1}^n {(p_{i-1} - o_i)}^2 } }
            RRSE = \sqrt{ \frac{ \sum_{i=1}^n {(p_i - o_i)}^2  }{ \sum_{i=1}^n {(p_{i-1} - o_i)}^2 }  }
        \end{equation}
\end{itemize}
where $o_i$ is the individual data $i = {1,...,n}$ and $p_i$ is the obtained prediction.

@@

Más concretamente, en un primer paso, el experto analizó las 50 variables de entrada que describen cada libro para eliminar algunas que estimó como suplerfuas.
Table \ref{tabla:params43} shows the set of variables describing a book (43 total inputs) that the expert manager determined as non-superfluous or non-redundant.

@@

Then, after analizyng the original dataset using the SOM to select the best variables to be considered, and taking into account the knowledge of the company experienced manager, 13 inputs were considered. 
Figure \ref{fig:som43} shows the initial analisys of dataset using the total of 43 descriptive inputs per book.
Table \ref{tabla:params13} shows the set of variables describing a book after the second analisys.
	% esta gráfica no aporta nada, ni la siguiente tampoco. No se puede
	% colocar en un trabajo en revista  cómo ha ido el entrenamiento, sino
	% datos serios de análisis estadístico de las variables. 
	% [Pedro] sí, estoy de acuerdo. Esto va con el ISSUE #6
\begin{figure}[!h] 
\begin{center}
\begin{tabular}{c}
  \epsfig{file=som43vars.eps,width=11cm} \\
  \epsfig{file=som43umatrix.eps,width=5cm}
\end{tabular}
\caption{\small{Initial data analysis, using 43 descriptive variables per book, using SOM.}}
	% Cojollos, ¿qué trabajo os cuesta poner esto en inglés directamente? - JJ
	% [Pedro] sí, si estoy de acuerdo... hay días que me "sale" el texto en inglés,
	% pero hay días que soltar la parrafada en español me resulta más fácil    :/
        % Sería más bien exploratory data analysis o data visualization. Un gráfico no es un análisis. Y pon las cosas cortas en inglés, hombre - JJ
\label{fig:som43}
\end{center}
\end{figure}

@@

Finally, after analizyng the dataset with 13 descriptive inputs per
book, and paying attention to the company experienced manager's
recommendations, 7 inputs were considered; specifically, variables
\emph{editorial} and \emph{materia} were included, while
\emph{devolucion\_total, reimpresiones, puntos\_vta1anio\_total,
  numero\_reimpresiones, devolucion1anio\_total, colocado1anio\_total,
  semanas\_venta, tirada} were discarded.  
	% Hay que aportar datos de por qué se hizo esto. ¿Análisis de planos?
	% ¿Análisis de varianza? - jj
	% [Pedro] En cuanto estén los SOM definitivos lo añadimos. Está relacionado con el ISSUE #6
Figure \ref{fig:som13} shows the initial analisys of dataset using the SOM on the total of 13 descriptive inputs per book.
This, along with the expert manager knowledge determined the final 7 inputs to be considered.
Table \ref{tabla:params7} shows the final set of variables describing a book, taking into account the company experienced manager knowledge.

@@

\begin{minipage}{2.3in} \begin{verbatim}
UNITS DISTRIBUTED IN FIRST EDITION
TOTAL SELLING POINTS
TOTAL DISTRIBUTED
RETAIL PRICE
SUBJECT-MATTER
EDITORIAL
TOTAL SALES (output)
\end{verbatim} \end{minipage}     \\

@@

Como se ha comentado anteriormente, existen diferentes métodos de predicción disponibles en la bibliografía, cada uno con una serie de parámetros diferente que pueden influir en los resultados.

Sin embargo, dado que en el caso de nuevos libros sólo ciertos datos descriptivos están disponibles (frente a datos históricos de ventas de libros ya publicados), no todos los métodos predictivos se pueden usar en este problema concreto.

Thus, in this research five forecasting methods, based on the literature review in the introduction \cite{Madsen2008,ChingChin2010,Thomassey2012,Xia2012,SThomassey2014}, have been used. 
Well known forecasting methods implemented in the Weka tool
\cite{Hall2009,Witten2011} have been choosen, as these methods are
widely known, and could be very helpful for practitioners to reproduce
experiments and even to solve similar sales prediction problems using
these methods. % eliminado el pie de página, es mejor usar referencias
               % - JJ

@@

Una vez que se han determinado los métodos a utilizar y los parámetros de ejecución, se aplicarán a los conjuntos de datos para obtener los modelos de predicción y posteriormente se comprobará su bondad ante datos no vistos previamente.

@@

  %[pedro] ya están los resultados de de tabla:results7 para MP5, LR, KNN, SVM.
  % faltan AUN los de MLP

  Obtained results using the different forecasting methods on the dataset with 7 variables per book can be shown in Table \ref{tabla:results7}:
\begin{table*}[!h]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
 Method   & Time (sec.)     & MAE       & RMSE    & RAE     & RRSE      \\
\hline
  MP5     & 7.39 $\pm$ 1.75   & 117.78   & 260.05  & 18.51   & 17.00  \\
\hline
  LR      & 40.59 $\pm$ 12.94 & 120.29  & 252.91  & 18.91   & 16.54   \\
\hline
  SVM     & 1547.91 $\pm$ 365.55 & 109.30  & 251.16   & 17.18     & 16.42   \\
\hline
  KNN     & 0.02 $\pm$ 0.01   & 293.98  & 1071.52  & 46.21   & 70.06   \\
\hline
  MLP     & 3265.71 $\pm$ 813.37   & 992.97 $\pm$ 0   & 1755.62 $\pm$ 0  & 156.07 $\pm$ 0   & 114.79 $\pm$ 0   \\
\hline
\end{tabular}
\end{center}
\caption{\small{Results (errors and time) obtained using the different forecasting methods on the dataset with 7 variables per book.}}
\label{tabla:results7}
\end{table*}

>>>>

In the case of already published books, available historical data might help 
to generate predictive models.
However, in the case of new books there is no historical data to define patterns, 
and only some descriptive data about the new book is available.
In these cases not any forecasting method can be used\cite{Madsen2008,ChernWSF15}.

@@

Throughout this research we collaborated with the company Trevenque Editorial\footnote{{\tt http://editorial.trevenque.es}}, 
a company that provides management systems and services for libraries, publishers 
and distributors.

Trevenque company, whose expert manager advised this research, provided 
the initial data set consisting of 3159 books.
In this data set, descriptive information for each book includes price, 
selling points, matter, publisher, sales, returns, among other information.
Table \ref{tabla:paramsOrig50} shows the original set of parameters describing 
each book.

@@

\begin{minipage}{2.3in} \begin{verbatim}
REFERENCE
AUTHOR
RETAIL PRICE
MATTER1
MATTER2
MATTER3
EDITORIAL
COLLECTION
BOOKBINDING
PRINT-RUN
TOTAL SALES
MALLS SALES
DELEGATES SALES
REST OF SALES
TOTAL SALES FIRST YEAR
MALL SALES FIRST YEAR
DELEGATES SALES FIRST YEAR
REST OF SALES FIRST YEAR
TOTAL DISTRIBUTED
MALLS DISTRIBUTED
DELEGATES DISTRIBUTED
REST OF DISTRIBUTED
TOTAL DISTRIBUTED FIRST YEAR
MALLS DISTRIBUTED FIRST YEAR
DELEGATES DISTRIBUTED FIRST YEAR
\end{verbatim} \end{minipage}     &

@@

In this prediction problem, several variables related to the own book affect 
the forecasts, such as author, price, matter or publisher.

However, there are other external variables that also affect the sales of 
a new book, such as those related to seasonality, trends or promotions, 
that classical prediction methods can not deal with.
However, an expert can consider them to modify the forecast obtained by 
the automatic method. Even publishing companies tend to define rules or 
indexes to subjectively adjust the results of prediction based on external data.

@@

There are several issues inherent to the problem of predicting sales of new 
products: the limitation on the amount of data available, the application of 
different methods and choose the most suitable of them.
In addition, there is no a standard methodology to solve the general problem 
of sales forecasting.
Usually, a preliminary analysis of the data is carried out to eliminate those 
variables that describe the product but do not provide useful information during 
the prediction process.

@@

Thus, in this study there are three main stages (see Figure \ref{fig:stages} 
for details): 
\begin{itemize}

  \item  First, after the editorial company has provided the raw data, taken 
  directly from its database, a selection variables process is carried out. 
  In this stage, an expert sales manager, using the information obtained 
  analyzing the SOM cluster method results, removes some unrelated variables 
  and selects the best variables to be considered in the forecasted model. 
  This task may help to identify leading indicators to be used by the 
  forecasting methods in order to improve the accuracy of future forecasts.

@@

  \item \emph{Root relative squared error} (RRSE):
        \begin{equation}\label{eq:RRSE}
            RRSE = \sqrt{ \frac{ \sum_{i=1}^n {(p_i - o_i)}^2  }{ \sum_{i=1}^n {(p_{i-1} - o_i)}^2 }  }
        \end{equation}
\end{itemize}
where $o_i$ is the individual data $i = {1,...,n}$ and $p_i$ is the obtained prediction.

@@

More specifically, in a first step, the expert analyzed the 50 input variables 
describing each book to remove those that were suplerfuas.
Table \ref{tabla:params43} shows the set of variables describing a book 
(43 total inputs) that the expert manager determined as non-superfluous or 
non-redundant.

@@

Then, after analizyng the original dataset using the SOM to select the best 
variables to be considered, and taking into account the knowledge of the company 
experienced manager, 13 inputs were considered. 
Figure \ref{fig:som43} shows the initial analisys of dataset using the total 
of 43 descriptive inputs per book.
Table \ref{tabla:params13} shows the set of variables describing a book after 
the second analisys.
	% esta gráfica no aporta nada, ni la siguiente tampoco. No se puede
	% colocar en un trabajo en revista  cómo ha ido el entrenamiento, sino
	% datos serios de análisis estadístico de las variables. 
	% [Pedro] sí, estoy de acuerdo. Esto va con el ISSUE #6
\begin{figure}[!h] 
\begin{center}
\begin{tabular}{c}
  \epsfig{file=som43vars.eps,width=11cm} \\
  \epsfig{file=som43umatrix.eps,width=5cm}
\end{tabular}
\caption{\small{Initial data analysis, using 43 descriptive variables per book, 
using SOM.}}
	% Cojollos, ¿qué trabajo os cuesta poner esto en inglés directamente? - JJ
	% [Pedro] sí, si estoy de acuerdo... hay días que me "sale" el texto en inglés,
	% pero hay días que soltar la parrafada en español me resulta más fácil    :/
        % Sería más bien exploratory data analysis o data visualization. 
        % Un gráfico no es un análisis. Y pon las cosas cortas en inglés, hombre - JJ
\label{fig:som43}
\end{center}
\end{figure}

@@

Finally, after analizyng the dataset with 13 descriptive inputs per
book, and paying attention to the company experienced manager's
recommendations, 7 inputs were considered; specifically, variables
\emph{EDITORIAL} and \emph{MATTER} were included, while
\emph{TOTAL DEVOLUTIONS, REPRINTS, TOTAL SELLING POINTS FIRST YEAR,
  NUMBER OF REPRINTS, TOTAL DEVOLUTIONS FIRT YEAR, TOTAL DISTRIBUTED FIRST YEAR,
  WEEKS ON SALE, PRINT-RUN} were discarded.  
	% Hay que aportar datos de por qué se hizo esto. ¿Análisis de planos?
	% ¿Análisis de varianza? - jj
	% [Pedro] En cuanto estén los SOM definitivos lo añadimos. Está relacionado con el ISSUE #6
Figure \ref{fig:som13} shows the initial analisys of dataset using the SOM on 
the total of 13 descriptive inputs per book.
This, along with the expert manager knowledge determined the final 7 inputs 
to be considered.
Table \ref{tabla:params7} shows the final set of variables describing a book, 
taking into account the company experienced manager knowledge.

@@

\begin{minipage}{2.3in} \begin{verbatim}
UNITS DISTRIBUTED IN FIRST EDITION
TOTAL SELLING POINTS
TOTAL DISTRIBUTED
RETAIL PRICE
MATTER
EDITORIAL
TOTAL SALES (output)
\end{verbatim} \end{minipage}     \\

@@

As previously stated, there are many prediction methods available in the 
literature, each one with a different parameter set that may affect the 
obtained results.

However, as in the case of new books only some descriptive data is available 
(compared to those cases in which historical data on sales of published books 
is available), not all forecasting methods can be used in this specific problem.

Thus, in this research five forecasting methods, based on the literature 
review in the introduction 
\cite{Madsen2008,ChingChin2010,Thomassey2012,Xia2012,SThomassey2014}, have been used. 
Well known forecasting methods implemented in the Weka tool
\cite{Hall2009,Witten2011} have been choosen, as these methods are
widely known, and could be very helpful for practitioners to reproduce
experiments and even to solve similar sales prediction problems using
these methods. % eliminado el pie de página, es mejor usar referencias - JJ

@@

Once  the methods to be applied have determined as well as their execution 
parameters, they must be applied to data sets to obtain the prediction models 
for each one. Later on, the models ability must be validated on non-previously 
seen data.

@@

  %[pedro] ya están los resultados de de tabla:results7 para MP5, LR, KNN, SVM.
  % faltan AUN los de MLP
Obtained results using the different forecasting methods on the dataset with 7 
variables per book can be shown in Table \ref{tabla:results7}:
\begin{table*}[!h]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
 Method   & Time (sec.)     & MAE       & RMSE    & RAE     & RRSE      \\
\hline
  MP5     & 7.39 $\pm$ 1.75   & 117.78   & 260.05  & 18.51   & 17.00  \\
\hline
  LR      & 40.59 $\pm$ 12.94 & 120.29  & 252.91  & 18.91   & 16.54   \\
\hline
  SVM     & 1547.91 $\pm$ 365.55 & 109.30  & 251.16   & 17.18     & 16.42   \\
\hline
  KNN     & 0.02 $\pm$ 0.01   & 293.98  & 1071.52  & 46.21   & 70.06   \\
\hline
  MLP     & 3265.71 $\pm$ 813.37   & 992.97 $\pm$ 0   & 1755.62 $\pm$ 0  & 156.07 $\pm$ 0   & 114.79 $\pm$ 0   \\
\hline
\end{tabular}
\end{center}
\caption{\small{Results (errors and time) obtained using the different 
forecasting methods on the dataset with 7 variables per book.}}
\label{tabla:results7}
\end{table*}
